Raising the Role of Vocabulary Hubs for Semantic Data Interoperability in Dataspaces

### 2024-06-12 EBDVF-2024/paper.qmd
add architecture description

## Architecture and Implementation

>>>>

## Software Architecture

The following figure shows the architecture of our approach:

![](img/architecture.png)

- A Data Provider sends data through the Connector, 
  and at the same time provides basic metadata (DCAT)
- Selected data can be stored in Database(s) managed by the dataspace,
  for easier access and querying.
  - For large-volume time series data, we recommend to use InfluxDB, PostgreSQL, Google BigQuery, etc.
  - Of course, access control is very important here.
    We'll leverage GraphDB's [Fine-grained Access Control](https://graphdb.ontotext.com/documentation/10.6/fine-grained-access-control.html), 
    but access control in the time series database is also important
  - Depending on Provider preferences and agreements, 
    other data may stay at the source, 
    and be delivered to the Consumer only dynamically through the Connector
- The Vocabulary Hub stores relevant semantic assets: ontologies and thesauri.
  These are added based on relevant use cases and datasets, following a dataspace governance process.
- The Vocabulary Hub also stores mapping assets, such as OBDA or R2RML mappings and CSVW manifests.
- PoolParty Semantic Integrator is used to edit, view and manage the semantic assets,
  and to bind mapping assets to incoming data.
- For virtual access to relational data, ONTOP is used to translate SPARQL to SQL dynamically
  and to convert returned data to SPARQL result format.
- The Data Consumer benefits from:
  - More powerful dataset discovery by using richer semantic queries, eg
    - "Give me all time series related to compressor C123" (equipment instance)
    - "Give me all time series related to temperature" (measured quantity)
    - "Give me all time series of bearings" (equipment kind or machine part)
  - Harmonized querying of datasets from multiple providers.
